---
title: "Skills Test 1"
output:
  rmdformats::html_docco
---

Reading in the data

```{r message=FALSE, warning=FALSE}
library(tidyverse)

data <- read.csv("https://raw.githubusercontent.com/mattols/geospat_data/main/earth-jobs.csv")
```


#1
List age and gender of the highest and lowest paid employees
```{r}
data[which.max(data$salary),c(2:3,7)]
data[which.min(data$salary),c(2:3,7)]
```


#2 
ordering dataset by salary in descending order

```{r}
data_ordered <- data[order(data$salary,decreasing = T),]
print(data_ordered)
```

#3
List the employees with a salary greater than $80000 whose hire data is before 2020-01-01

```{r}
data_ordered$hire_date <- as.Date(data_ordered$hire_date)

data_ordered[data_ordered$salary > 80000 & data_ordered$hire_date < "2020-01-01",]
```

#4

```{r}
med_geo <- median(data_ordered[data_ordered$area == "Geology",]$salary)

med_envt <- median(data_ordered[data_ordered$area != "Geology",]$salary,na.rm = T)

#difference between median
med_geo-med_envt
```

#5

```{r}
data_ordered$date_diff <- Sys.Date() - data_ordered$hire_date

data_ordered$year_diff <- as.numeric(data_ordered$date_diff / 365.25)

print(data_ordered)
```

#6

```{r warning=FALSE}
data_ordered %>% 
  ggplot(aes(x=age,y=salary,col=gender)) +
  geom_point()+
  facet_wrap(~area) +
  theme_bw()
```

Granted this is a smaller data set but it almost seems like there is a negative correlation between age and salary, especially for environmental science. While it does seem like being male you would be making more based on this graph, there is also simply fewer female data points to base on average off of (which could be its own problem)


```{r warning=FALSE}
data_ordered %>% 
  ggplot(aes(x=education,y=salary,fill=gender)) +
  geom_col(position = "dodge") +
  facet_wrap(~job) +
  theme_bw()
```

To get the highest pay by education you should aim at getting a Masters, this makes significantly more than any of the other areas, including a PhD. While there are less female data points to base on average off of, it does seem that men with the same education typically make a couple grand more a year. Except for data science jobs, this dataset shows women making more with the same education and field.

#7

```{r}
summary(lm(salary ~education + job + area + gender + age,data_ordered))
```

It seems like the majority of these variables can help in predicting salary, given a person with all of these variables together you could get a pretty good guess of their salary based on the p-value of 4.8E-7 and the R^2 of 0.9115

#8

```{r}
vect <- sample(100)

#Creating sample variance function
sam_var <- function(vect) {
  #creating total variable
  total = 0
  
  #calcuating the total sum of the vector
  for (i in 1:length(vect)) {
    total <- total + vect[i]
  }

  #calculating the mean
  v_mean <- total / length(vect)
  
  #creating total square distance variable
  tot_sq_dist = 0
  
  #loop to calculate total square distance
  for (i in 1:length(vect)) {
    sq_dist <- (vect[i]-v_mean)^2
    tot_sq_dist = tot_sq_dist + sq_dist
  }
  sample_variance <- tot_sq_dist/(length(vect)-1)
  
  return(sample_variance)
}

sam_var(vect)

```

