---
title: "Task 3"
output: html_document
date: "2024-02-02"
---

# 1.
#### 1.1
##### Find the number of days in the semester.
```{r}
start_date <-  as.Date("01/08/2024",tryFormats = "%m/%d/%Y")
end_date <- as.Date("04/23/2024",tryFormats = "%m/%d/%Y")
time_frame <- end_date - start_date
print(time_frame)
```
#### 1.2
##### Finding how many saturdays are in the semester

```{r}
#Extract all dates that fall on a Saturday and count the total number of Saturdays in the semester.
all_dates <- seq(start_date, end_date, by = "days")

# Extract all dates that fall on a Saturday
saturdays <- all_dates[weekdays(all_dates) == "Saturday"]
```

#### 1.3
##### Finding how many hours we meet in the semester
```{r}
meeting_days <- all_dates[weekdays(all_dates) == "Tuesday" | weekdays(all_dates) == "Thursday"]

#This is how many hours we will meet throughout the semester
length(meeting_days) * 3
```

#### 1.4
##### Extracting valid emails

```{r}
# emails
emails <- c("w2u9q0g7z6@outlook.com", "m7n4w6f1a0@outlook.com", "withspaces@gmail.com",  
              "n7y4h5k0t6@gmail.com", "@invalid", "e2m5h7q1z0@outlook.com", 
              "invalid@", "invalid@", "invalid@",
              "g2b1j5u7o8@outlook.com", "s1c8k0t6e7@gmail.com", "f9m1a4z8h6@outlook.com",
              "e8c5f3v6k9@yahoo.com", "missingdotgmail.com", "invalid@email",
              "b6g3f0y4d2@gmail.com", "k0p8i9s6o3@outlook.com", "s8t7g2k9o5@yahoo.com", 
              "a9i7y1w5q0@gmail.com", "r2g9y4l5i6@yahoo.com", "b7d4a5u0j2@outlook.com",
              "l0a7z9m6r1@outlook.com", "c4v2a7w6r1@yahoo.com", "x1s3d8y5c6@yahoo.com",
              "@invalid", "@invalid", "@invalid",
              "@invalid", "invalid@", "e3k9x2m8t4@gmail.com",
              "invalidemail", "k2a3b7l1n8@outlook.com", "d0a6q9u2z4@yahoo.com",
              "i0f2z8c5x1@outlook.com", "q4m0l5o6j2@outlook.com", "v5h9a4f1q0@outlook.com",
              "@invalid", "f0z3h5d1v2@gmail.com", "g7k0u3a5s6@gmail.com",
              "@invalid", "invalid@", "@invalid", 
              "@invalid", "n9t0h3e6i5@outlook.com", "l2k1r6z0g5@outlook.com",
              "f9d7u5e1n8@outlook.com", "n9o4s8a6b0@yahoo.com", "@invalid",
              "s1b8q5i9z2@gmail.com", "z7v3n1o0f9@outlook.com")

#creating new vector with emails that contain the @ sign
emails_with_at_sign <- emails[grep("@",emails)]

#using the @ sign vector and keeping only the emails with .com
valid_emails <- emails_with_at_sign[grep(".com",emails_with_at_sign)]

valid_emails
```

#### 1.5
##### finding valid phone numbers 

```{r}
# phone numbers
phones <- c("(525) 692-8788", "(748) 654-2046", "(865) 970-8066", "(310) 472-4003", 
            "(689) 943-4206", "920.680.4936", "323.171.5088", "265.687.3906", 
            "316.674.1293", "389.240.9468", "562 322 3503", "140 627 7741", 
            "530 215 9943", "189 705 9565", "415 873 7128", "133-854-8126", 
            "615-508-4357", "112-407-4979", "168-377-4229", "994-188-6602",
            "123-456-789", "abc-123-4567", "123 45 6789", "1a2b3c4d5e", 
            "12345", "67890", "1234", "6789", "12345678903", "12345678901")

#cleaning up phone numbers
#removing hyphens
phones <- gsub("-","", phones)

#removing dots
phones <- gsub("\\.","",phones)

#removing parenthesis
phones <- gsub("[()]","",phones)

#removing spaces
phones <- gsub(" ","",phones)

#replacing any a,b,or c with 2
phones <- gsub("[abc]","2",phones)

#replacing any d,e,or f with 3
phones <- gsub("[def]","3",phones)

#replacing any g,h,or i with 4
phones <- gsub("[ghi]","4",phones)

#replacing any j,k or l with 5
phones <- gsub("[jkl]","5",phones)

#replacing any m,n,or o with 6
phones <- gsub("[mno]","6",phones)

#replacing any p,q,r,or s with 7
phones <- gsub("[pqrs]","7",phones)

#replacing any t,u,v with 8
phones <- gsub("[tuv]","8",phones)

#replacing any w,x,y,or z with 9
phones <- gsub("[wxyz]","9",phones)

#use nchar function which tells the number of characters for each element in the vector
character_lengths <- nchar(phones)

#subset to keep only elements with 10 characters
valid_phones <- phones[character_lengths == 10]

valid_phones
```

#### 1.6
##### Adding underscores in the phone numbers and checking if they are valid
```{r}
#adding underscores after the first 3 and the first 6 characters
valid_phones <- gsub("^(.{3})(.{3})", "\\1_\\2_", valid_phones)

valid_phones

#checking in phone numbers are valid

#making a pattern that is xxx_xxx_xxxx
pattern <- "^\\d{3}_\\d{3}_\\d{4}$"

#using grepl to see boolean values if they follow that pattern
grepl(pattern, valid_phones)
```

# 2.
#### 2.7
##### Reading in snotel data and making a statement for the data

```{r}
##### Part 2 #####
#2. Snotel data

# Read in csv file provided in canvas
snotel <- read.csv("./snotel820_timp.csv")

# After reading in the data, convert dates column to a date format
snotel$date <- as.Date(snotel$date)

#estimating the density of the snowpack each day
snotel$density = snotel$swe/snotel$depth * 100

# List the average and standard deviation of the following: 
#snow density (in percent)
mean_density <- round((mean(snotel$density)), 2)
sd_density <- round((sd(snotel$density)), 2)

#snow depth
mean_depth <- round((mean(snotel$depth)),2)
sd_depth <- round((sd(snotel$depth)),2)

#average temperature
mean_ave_temp <- round((mean(na.omit(snotel$t_ave))),2)
sd_ave_temp <- round((sd(na.omit(snotel$t_ave))),2)

#warmest and coldest temperature recorded during the time period
max_temp <- max(na.omit(snotel$t_max))
min_temp <- min(na.omit(snotel$t_min))

#Concatenate this information into a single print or cat statement
cat("Density Average = ", mean_density, "%",
                    "\n",
                    "Density Standard Deviation = ", sd_density, "%",
                    "\n",
                    "Depth Average = ", mean_depth, '"',
                    "\n",
                    "Depth Standard Deviation = ", sd_depth,'"',
                    "\n",
                    "Temperature Average = ", mean_ave_temp, "째F",
                    "\n",
                    "Temperature Standard Deviation = ", sd_ave_temp, "째F",
                    "\n",
                    "Max Temperature Recorded = ", max_temp, "째F",
                    "\n",
                    "Minimum Temperature Recorded = ", min_temp, "째F")
```


#### 2.8
##### plotting density over time and depth over time

```{r}
# Setting the larger right margin
#First value is the bottom line
#second value is the left line
#third value is the top line
#fourth value is the right line
par(mar = c(5, 5, 5, 5))

plot(snotel$date,
     snotel$depth,
     col = "blue",
     xlab = "Date",
     ylab = "Depth (inches)",
     ylim = c(min(snotel$depth), max(snotel$depth)),
     type = "l")

title(main = "Depth and Density Over Time")
axis(2, col.axis = "blue")

par(new = TRUE)

plot(snotel$date,
     snotel$density,
     col = "red",
     xlab = "",
     ylab = "",axes = FALSE,
     type = "l")

axis(4, col.axis = "red")
mtext("Density (percentage)", side = 4, line = 2.5, col = "red",)
```

**looking at the graph, jan 4th seems to be the day that marks the transition from clear to stormy weather**


#### 2.9

```{r}
#Is there a correlation between snow water equivalent and precipitation accumulation? Describe the strength and direction of any relationship.
plot(snotel$swe,
     snotel$prec_accum)

model <- lm(snotel$swe ~ snotel$prec_accum)

strength <- round(as.numeric(model$coefficients[2]),3)

paste0("Positive correlation of ", strength, " when comparing swe to precipitation accumulation")
```

#### 2.10

```{r}
# Create a new column called swe_perc showing the snow water equivalent as a percent of the median historical swe value for each day. Plot these values and describe how the swe values track with the median historical values?

snotel$swe_perc <- snotel$swe/snotel$swe_med * 100
plot(snotel$date,snotel$swe_perc)
```

#### 2.11

```{r}
#   How much snow (depth) was gained or lost from the start to end of time period? How much snow water equivalent was gained or lost?

max_day_df <- snotel[(snotel$date == max(snotel$date)),]
min_day_df <- snotel[(snotel$date == min(snotel$date)),]

max_day_df$depth - min_day_df$depth
max_day_df$swe - min_day_df$swe
```

#### 2.12
```{r}
#   Write a script to calculate the daily incremental change in both snow depth and SWE over the 30 day period. Sum both the gains and losses from each and compare these values to the values in the previous question. What might account for the differences?
differences = NULL

for (i in 1:(length(snotel$depth) - 1)) {
  diff = snotel$depth[i+1]-snotel$depth[i]
  print(diff)
  differences <- c(differences, diff)
}
```

This is how many inches were gained throughout the entire dataset

```{r}
sum(differences)
```



```{r}
differences2 = NULL

for (i in 1:(length(snotel$swe) - 1)) {
  diff2 = snotel$swe[i+1]-snotel$swe[i]
  print(diff2)
  differences2 <- c(differences2, diff2)
}
```

This is how many inches of snow water equivalent were gained throughout the dataset

```{r}
sum(differences2)
```
The difference is because there was snow gain AND snow loss throughout the duration of the entire dataset.

#### 2.13
##### Looking at snow depth vs temperature

```{r}
linear_model <- (lm(snotel$t_ave~snotel$depth))
plot(snotel$t_ave,snotel$depth)
abline(linear_model,col = "green")

summary(linear_model)
```
An R^2 of 0.01366 and a p-value of 0.546

```{r}
sub_snotel = snotel[snotel$date >= as.Date("2024/1/04"),]

linear_model2 <- lm(sub_snotel$t_ave~sub_snotel$depth)
plot(sub_snotel$t_ave,sub_snotel$depth)
abline(linear_model2,col = "green")

summary(linear_model2)
```
An R^2 of 0.3683 and a p-value of 0.007566.

The relationship is much more correlative now, this is because we took out the days where snow was actually melting which was making the data all messed up when doing an analysis on temperature and depth.

# 2.15

```{r}
linear_model <- (lm(snotel$t_ave~snotel$density))
plot(snotel$t_ave,snotel$density)
abline(linear_model,col = "green")

summary(linear_model)
```

While there isn't a great correlation between density and temperature visually, the p-value is okay, but the R^2 is pretty bad.

```{r}
linear_model <- (lm(snotel$depth~snotel$density))
plot(snotel$depth,snotel$density)
abline(linear_model,col = "green")

summary(linear_model)
```
There is a pretty good correlation between snow depth and density, this makes sense as you would expect it to get more dense the more snow depth as it is compressing the snow below it.
