---
title: "Task 8"
output: 
  html_document:
    code_folding: hide
---

### Packages required: **Terra** & **leaflet**

```{r message=FALSE, warning=FALSE, class.source = "fold-show"}
library(terra)
library(leaflet)
```

Read this article, it gives a good overview of remote sensing and why band calculations work. [NDVI, NDBI, NDWI Calculation Using Landsat 7/8](https://www.linkedin.com/pulse/ndvi-ndbi-ndwi-calculation-using-landsat-7-8-tek-bahadur-kshetri)

### Part 1 final product: **Create these two plots**

```{r echo=FALSE}
c_ls_7 <- rast("../../landsat/LE07_CU_008008_20010718_20210426_02_SR/cropped_landsat7.tif")

MNDWI <- (c_ls_7[[2]] - c_ls_7[[5]]) / (c_ls_7[[2]] + c_ls_7[[5]]) > 0
NDBI <- (c_ls_7[[5]] - c_ls_7[[4]]) / (c_ls_7[[5]] + c_ls_7[[4]]) > 0
NDVI <- (c_ls_7[[4]] - c_ls_7[[3]]) / (c_ls_7[[4]] + c_ls_7[[3]]) > 0.2

#For Landsat 7 data, MNDWI = (Band 2 – Band 5) / (Band 2 + Band 5)
plotRGB(c_ls_7, r = 3, g = 2, b = 1, main = "RGB Image over Wasatch Mountains", stretch = "lin", axes = T, grid = T, mar = c(2,2,2,2))

plot(MNDWI, col = c("transparent","cadetblue1"), main = "Classified image",  add = F, legend = F)
plot(NDVI, add = TRUE,col = c("transparent","darkgreen"), legend = F)
plot(NDBI, add = TRUE, col = c("transparent","tan"),legend = F) 

legend(x= 410000, y= 4500000, legend = c("Water", "Barren", "Vegetation"), fill = c("cadetblue1", "tan", "darkgreen"))
```

---

The Normalized Difference Vegetation Index (NDVI) is calculated as:

\[
NDVI = \frac{(NIR - Red)}{(NIR + Red)}
\]

where NIR represents the Near Infrared band **(Band 4)** and the Red band **(Band 3)** for Landsat 7.

---

The Modified Normalized Difference Water Index (MNDWI) is calculated as:

\[
MNDWI = \frac{(Green - SWIR)}{(Green + SWIR)}
\]

where Green represents the Green band **(Band 2)** and SWIR represents the Shortwave Infrared band **(Band 5)** of Landsat 7.

---

The Normalized Difference Built-Up Index (NDBI) is typically used to see areas of urban development, but it works well with this image to calculate barren land, it is calculated as:

\[
NDBI = \frac{(SWIR - NIR)}{(SWIR + NIR)}
\]

where SWIR represents the Shortwave Infrared band **(Band 5)** and NIR represents the Near Infrared band **(Band 4)** of Landsat 7.

---

### Steps:

-Download the landsat 7 image.

-Read in the image.

-Do band caluclations for MNDWI, NDBI, and NDVI

-Create threshold for those band calculations so they are true/false values

-Plot all three band calculation thresholds on the same image

-Add Legend with corresponding colors

---

*Hints:*

What if my RGB plot is really dark?
```{r class.source = "NULL"}
# If your plotRGB plot is really dark this is because the image needs to be stretch. There are a few different stretch methods:

# stretch = "lin"
# This stretch method applies a linear stretch of values across the image

# stretch = "hist"
# This stretch method applies a histogram stretch of values across the image
```

What if my classification map does not look like the expected product?
```{r class.source = "NULL"}
# This may not be an issue and is just due to your threshold being different than mine for each index. 
# These are the values I used
# MNDWI > 0
# NDBI > 0 
# NDVI > 0.2
```

What if I can't get the false values to be transparent?
```{r class.source = "NULL"}
# After creating a threshold for the indexes you can plot them with transparent false values like so in the plot function: col = c("transparent","cadetblue1") This sets the false value color first and then the true value color.
```

How do I access specific layers in a raster with multiple layers?
```{r, class.source = "NULL"}
# You can access a specific layer through indexing, it's a little different for raster you use double brackets. EX:  image[[x]] with x being the layer number you want to grab. Be careful assuming that each index number directly correlated to the band number, they may be out of order!
```

---

### Part 2 final product:

```{r class.source = "NULL"}
dem <- rast("../../DEMs/timp/output_USGS1m.tif")

c1 <- crop(dem,c(442000,448000,4470000,4474000))
c2 <- crop(dem,c(443000,449000,4467000,4471000))

c1 <- aggregate(c1,10)
c2 <- aggregate(c2,10)

aspect <- terrain(mosaic(c1,c2),"aspect")
slope <- terrain(mosaic(c1,c2),"slope")

mos <- mosaic(c1,c2)

leaflet()%>%
  addProviderTiles("OpenStreetMap.HOT") %>%
  addRasterImage(shade(slope*pi/180, aspect*pi/180, angle = 40, direction = 270), col = grey(0:100/100)) %>% 
  addMiniMap()
```


---

### Steps:

-Download the DEMs

-Mosaic the DEMs together

-Create an aspect and slope map

-Create a hillshade map with the aspect and slope map

-Plot a leaflet map with a few different options (minimap, bounding box, etc...)

---

Copy this and figure out how to add the mosaic of the DEM. Add a few different elements like a minimap and try to add some fun stuff!

The way that I've gone about discovering the possibilities of the leaflet package is by typing **leaflet::** and seeing all of the functions that it has!

```{r eval = F, class.source = "fold-show"}
leaflet()%>%
  addProviderTiles("OpenStreetMap.HOT")
```


---

*Hints:*

How do I make a hillshade from a slope and aspect map?
```{r class.source = "NULL"}
# shade(slope*pi/180, aspect*pi/180, angle = 40, direction = 270)

# col = grey(0:100/100)
```



