---
title: "Quiz 3"
author: "Noah Christensen"
output: rmdformats::downcute
---

___

Here's the packages needed to perform the functions throughout this document:

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(broom)
```

# Part 1


## Cleaning

Let's read in the data and clean it up

```{r}
#reading in dataframe
df <- read.csv("./FacultySalaries_1995.csv")

#cleaning names
df <- df %>% clean_names()

#pivoting longer all columns that end with salary
df <- df %>% pivot_longer(ends_with("salary"),
                    names_to = "rank",
                    values_to = "salary")

#getting rid of "VIIB" in tiers
df <- df %>% filter(tier != "VIIB")

# Extract text between the first and second underscore
df$rank <- str_extract(df$rank, "(?<=_)[^_]+(?=_.*)")
```

___

## Plotting


Now the dataframe is cleaned let's get plotting

```{r warning=FALSE}
df %>% 
  ggplot(aes(x=rank,y=salary,fill=factor(rank))) +
  geom_boxplot() +
  facet_grid(~tier) +
  labs(y="Salary",
       x="Rank",
       fill="Rank") +
  theme_bw() +
  scale_fill_discrete(labels = c("Assist", "Assoc", "Full")) +
  scale_x_discrete(labels = c("Assist", "Assoc", "Full"))
```

___

## Modeling

Creating an anova model

```{r}
#building and displaying anova model
amod <- aov(data= df, salary ~ rank + state + tier)

#viewing model
summary(amod)
```

# Part 2


## Cleaning


Let's load in the new data set and clean it up

```{r}
#reading in and cleaning names
df2 <- read.csv("./Juniper_Oils.csv") %>% 
  clean_names()

#pivoting longer
df2 <- df2 %>% 
  pivot_longer(cols = c("alpha_pinene","para_cymene","alpha_terpineol","cedr_9_ene","alpha_cedrene","beta_cedrene","cis_thujopsene","alpha_himachalene","beta_chamigrene","cuparene","compound_1","alpha_chamigrene","widdrol","cedrol","beta_acorenol","alpha_acorenol","gamma_eudesmol","beta_eudesmol","alpha_eudesmol","cedr_8_en_13_ol","cedr_8_en_15_ol","compound_2","thujopsenal"),
               names_to = "chemicalID",
               values_to = "concentration")
```

___

## Plotting


Plotting data

```{r message=FALSE}
df2 %>% 
  ggplot(aes(x=years_since_burn,y=concentration)) +
  geom_smooth() +
  facet_wrap(~chemicalID, scales = "free") +
  theme_bw()
```

___

## Modeling


Creating a model that tells which chemicals have signification concentration changes by the amount of years since the burn 

```{r}
# creating model
mod2 <- glm(data=df2, formula = concentration ~ years_since_burn * chemicalID)

# Get tidy output
tidy_output <- tidy(mod2)

# Filter for significant results (adjust the threshold as needed)
significant_results <- subset(tidy_output, p.value < 0.05)

# renaming term column
updated_results <- significant_results %>%
  mutate(term = sub("^years_since_burn:","",term)) %>% 
  mutate(term = sub("^chemicalID", "", term))
  
# View the modified data frame
updated_results
```

You can see that there are 6 different chemicals that have significant effect on concentration, given the threshold of a p value less than 0.05, with cedrol having the strongest correlation.
